---
title: "Quick Start: Current causalBKMR API"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{gbkmr_quickstart}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, purl=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
```

This vignette uses the current package implementation in `R/01-validation.R`
through `R/05-user-interface.R`. The archived code in `R/old/` is not
part of this workflow.

The current user-facing path is:

1. Put the outcome, mixture history, and covariate history into matrices.
2. Convert those matrices with `prepare_gbkmr_data()`.
3. Check the detected variable structure.
4. Run `gbkmr_run()`.

## Input Contract

For `T` time points, `p` mixture components, `Ldim` time-varying
covariates, and `B` baseline covariates, the current API expects:

| Object | Shape | Ordering |
|--------|-------|----------|
| `Y` | `n` vector | final outcome |
| `Z` | `n x (p * T)` matrix | `M1_t0, M2_t0, ..., M1_t1, ...` |
| `X` | `n x (Ldim * (T - 1) + B)` matrix | post-baseline time-varying covariate columns first, then baseline columns |

`prepare_gbkmr_data()` writes mixture columns as `logM1_0`, `logM2_0`,
..., time-varying covariates as `<name>_1`, `<name>_2`, ..., and
baseline covariates as `baseline_1`, `baseline_2`, ....

## Simulate Data

```{r simulate, purl=FALSE}
library(causalBKMR)

set.seed(123)
n <- 300
T_points <- 3
p <- 2
Ldim <- 1
B <- 1

baseline_1 <- rbinom(n, 1, 0.5)
waist_base <- rnorm(n, 85, 12)

Z <- matrix(rlnorm(n * p * T_points, meanlog = 1, sdlog = 0.5),
            nrow = n, ncol = p * T_points)

waist_t1 <- waist_base + 0.4 * log(Z[, 1]) + rnorm(n, 0, 2)
waist_t2 <- waist_t1 + 0.3 * log(Z[, 3]) + rnorm(n, 0, 2)
X <- cbind(waist_t1, waist_t2, baseline_1)

Y <- 50 +
  0.5 * baseline_1 +
  0.15 * waist_t2 +
  0.8 * log(Z[, 5]) -
  0.4 * log(Z[, 6]) +
  rnorm(n, 0, 4)
```

## Prepare Data

```{r prepare, purl=FALSE}
prepared <- prepare_gbkmr_data(
  Y = Y,
  Z = Z,
  X = X,
  time_points = T_points,
  mixture_components = p,
  td_covariates = Ldim,
  baseline_covariates = B,
  td_covariate_names = "waist",
  log_transform_mixtures = TRUE
)

names(prepared)
head(prepared)
```

Expected columns include:

```text
baseline_1
logM1_0 logM2_0 logM1_1 logM2_1 logM1_2 logM2_2
waist_1 waist_2
Y
id
```

## Check Detection

```{r detect, purl=FALSE}
detection <- detect_variable_patterns(prepared, T = T_points)
print(detection)
```

This check matters because `gbkmr_run()` uses the detected structure to
build models for time-varying confounders and outcomes. If `p`, `Ldim`, or the covariate
names are wrong, fix the matrices before fitting.

## Run Analysis

For real analyses use many more iterations. The small values below are
only a smoke-test template.

```{r run, purl=FALSE}
fit <- gbkmr_run(
  data = prepared,
  outcome = "Y",
  outcome_type = "continuous",
  time_points = T_points,
  engine = "auto",
  n = nrow(prepared),
  iter = 15000,
  K = 1000,
  n_knots = 50
)
```

`n_knots = 50` uses the predictive-process approximation to the Gaussian
process. It is faster, but it narrows the posterior intervals and lowers their
coverage; `n_knots = NULL` fits the exact Gaussian process as in Chai et al.
With `engine = "bkmr"`, `n_cores` forks that many workers for the model fits
and the Monte Carlo g-computation (`parallel::mclapply`, not available on
Windows); results do not depend on `n_cores`.

`engine = "auto"` selects standard BKMR for `n <= 2000`, binary
outcomes, or binary time-varying confounders. For larger all-Gaussian
analyses it uses `fastbkmr` only if the `fbkmr` package is installed.

## Read Results

```{r results, purl=FALSE}
print(fit)
summary(fit)

fit$causal_effect
fit$counterfactual_means
fit$detection_info
fit$diagnostics
```

The default estimand is the average causal effect of shifting every
mixture component at every time point from its 25th percentile to its
75th percentile:

```text
ACE = E[Y(a*)] - E[Y(a)]
```

Only mixture columns are intervened on. Time-varying confounders are
sampled forward through fitted time-varying confounder models, and baseline
covariates enter the BKMR models linearly.

## Next Steps

- [Methodology overview](gbkmr_method_overview.html)
- [Custom intervention contrasts](gbkmr_example_contrasts.html)
- [Binary outcome](gbkmr_example_binary.html)
- [Large samples with fastBKMR](gbkmr_example_fastbkmr.html)
- [Diagnostics and troubleshooting](gbkmr_diagnostics.html)
