---
title: "Large Samples with fastBKMR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{gbkmr_example_fastbkmr}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, purl=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
```

This vignette describes how the current `gbkmr_run()` interface uses the
standard BKMR engine and the optional `fastbkmr` engine. The data
preparation and intervention arguments are identical across engines.

Archived functions under `R/old/` are not used.

## Engine Choice

Version 1 separates the standard BKMR path from the fastBKMR path:

| Outcome | Time-dependent confounder `L` | `engine = "bkmr"` | `engine = "fastbkmr"` |
|---------|-------------------------------|-------------------|-----------------------|
| continuous | all continuous | Supported: Gaussian outcome + Gaussian `L` models | Supported: Gaussian fastBKMR throughout |
| binary | all continuous | Supported: probit outcome + Gaussian `L` models | Not supported; hard error |
| continuous | any binary | Supported: Gaussian outcome + probit models for binary time-varying confounders with Bernoulli Monte Carlo sampling | Not supported; hard error |
| binary | any binary | Supported: probit outcome + probit models for binary time-varying confounders with Bernoulli Monte Carlo sampling | Not supported; hard error |

Here, binary means a 0/1 outcome or a 0/1 time-dependent confounder.
Mixture exposures can still be continuous. `fbkmr::skmbayes()` in the
current public `fbkmr` package is a Gaussian fast path; generalized or
logistic fastBKMR is not yet available in this installed package.

With `engine = "auto"`, `gbkmr_run()` chooses `fastbkmr` only for large
analyses with a continuous outcome and all continuous time-dependent
confounders when `fbkmr` is installed. Otherwise it uses standard `bkmr`.

## Install the Optional Engine

`fbkmr` is not a required dependency of `causalBKMR`. Install it only if
you plan to run large continuous-outcome analyses.

```{r install-fbkmr, purl=FALSE}
install.packages("remotes")
remotes::install_github("junwei-lu/fbkmr")
```

Parallel fastBKMR runs also use `doSNOW` and `doParallel` when available.

```{r install-parallel, purl=FALSE}
install.packages(c("doSNOW", "doParallel"))
```

## Automatic Selection

Use the same prepared data object created by `prepare_gbkmr_data()`:

```{r auto, purl=FALSE}
library(causalBKMR)

fit_auto <- gbkmr_run(
  data = prepared_large,
  outcome = "Y",
  outcome_type = "continuous",
  time_points = 3,
  engine = "auto",
  n = nrow(prepared_large),
  iter = 15000,
  K = 1000
)
```

For `nrow(prepared_large) > 2000`, this continuous-outcome example uses
`fastbkmr` if `fbkmr` is installed and all time-dependent confounders are
continuous. The `n` argument is optional when
`data` is supplied, but setting it explicitly makes the intended sample
size clear in reports and scripts.

## Force fastBKMR

You can also request `fastbkmr` directly:

```{r force-fast, purl=FALSE}
fit_fast <- gbkmr_run(
  data = prepared_large,
  outcome = "Y",
  outcome_type = "continuous",
  time_points = 3,
  engine = "fastbkmr",
  n_subset = 6,
  n_cores = 6,
  iter = 15000,
  K = 1000
)
```

`n_subset` controls how many subsets are fit. `n_cores` controls how many
parallel workers are requested. If parallel execution is unavailable,
the implementation falls back to a sequential run. (With `engine = "bkmr"`,
`n_cores` instead forks workers for the BKMR fits and the Monte Carlo
g-computation.)

## Compare Engines on a Subset

For a new analysis, it is useful to compare the engines on a smaller
subset before launching a large run.

```{r compare-engines, purl=FALSE}
set.seed(1)
idx <- sample(seq_len(nrow(prepared_large)), 1500)
prepared_subset <- prepared_large[idx, ]

fit_bkmr <- gbkmr_run(
  data = prepared_subset,
  time_points = 3,
  engine = "bkmr",
  iter = 15000,
  K = 1000,
  n_knots = 50
)

fit_fast_subset <- gbkmr_run(
  data = prepared_subset,
  time_points = 3,
  engine = "fastbkmr",
  n_subset = 5,
  n_cores = 5,
  iter = 15000,
  K = 1000
)

fit_bkmr$causal_effect
fit_fast_subset$causal_effect
```

Small differences are expected because the engines use different
posterior approximations. Large differences are a signal to inspect
diagnostics, intervention support, and model settings.

## Limitations

- Binary outcomes and binary time-dependent confounders require
  `engine = "bkmr"`.
- `n_knots` applies to standard BKMR only; `n_knots = NULL` fits the exact
  Gaussian process (wider, better calibrated intervals than the knot
  approximation).
- `fastbkmr` is optional; `engine = "auto"` falls back to `bkmr` when
  `fbkmr` is not installed.
- The raw fit object returned by `fit$raw_results` differs by engine:
  standard BKMR stores one fit per model, while fastBKMR stores subset
  fits.

## See Also

- [Quick Start](gbkmr_quickstart.html)
- [Binary outcomes](gbkmr_example_binary.html)
- [Diagnostics and troubleshooting](gbkmr_diagnostics.html)
